---
title: "Exercise 8: Summary Tables with gtsummary"
format: html
editor_options: 
  chunk_output_type: console
---


In this exercise, you will practice using the `gtsummary` package to create summary tables using the low birth weight dataset.

### **Instructions**

You are provided with the low birth weight dataset, which includes variables such as birth weight (`bwt`), mother's weight (`lwt`), mother's age (`age`), smoking status (`smoke`), and race (`race`). Follow the questions below to analyze the dataset using the `gtsummary` package.

### **Question 1:** Load the Data and Packages

- Load the dataset using `read_csv()` and `here()` functions.
- Load the necessary packages using `pacman::p_load()`.

```{r}
#| echo: false
#| warning: false
#| message: false

# Load required packages and data
pacman::p_load(tidyverse, here, gtsummary)

df <- read_csv(here("data", "low_birth_weight.csv"))
```


### **Question 2:** Create a Summary Table for Continuous Variables

- Generate a summary table for the following continuous variables:
  - Birth weight (`bwt`)
  - Mother's weight (`lwt`)
  - Mother's age (`age`)
  
- Use the `tbl_summary()` function to display the median and interquartile range (IQR) for each variable.

- Use appropriate labels for each variable

```{r}
#| echo: false


# Question 2: Summary Table for Continuous Variables

df |> 
  dplyr::select(age, lwt, bwt) |>
  tbl_summary(
  type = list(bwt ~ "continuous", lwt ~ "continuous", age ~ "continuous"),
  statistic = all_continuous() ~ "{median} ({IQR})",
  label = list(age ~ "Age of Mother", 
               lwt ~ "Weight of Mother",
               bwt ~ "Birth Weight of the Child")
)
```

#### Answer

```{r}
#| code-fold: true
#| eval: false


# Question 2: Summary Table for Continuous Variables

df |> 
  dplyr::select(age, lwt, bwt) |>
  tbl_summary(
  type = list(bwt ~ "continuous", lwt ~ "continuous", age ~ "continuous"),
  statistic = all_continuous() ~ "{median} ({IQR})",
  label = list(age ~ "Age of Mother", 
               lwt ~ "Weight of Mother",
               bwt ~ "Birth Weight of the Child")
)
```

### **Question 3:** Create a Summary Table for Categorical Variables

- Generate a summary table for the following categorical variables:
  - Smoking status (`smoke`)
  - Race (`race`)
  - Low birth weight (`low`)
  
- Use appropriate labels for each variable and include the counts and percentages.

```{r}
#| echo: false

df |> 
  select(smoke, race, low) |> 
gtsummary::tbl_summary(
  label = list(smoke ~ "Smoking Status", race ~ "Race", low ~ "Low Birth Weight"),
)
```


#### Answer

```{r}
#| code-fold: true
#| eval: false


df |> 
  select(smoke, race, low) |> 
gtsummary::tbl_summary(
  label = list(smoke ~ "Smoking Status", race ~ "Race", low ~ "Low Birth Weight"),
)
```

### **Question 4:** Stratified Summary Table by Low Birth Weight Status

- Create a summary table for continuous variables stratified by the low birth weight status (`low`).
- Display the median and IQR for each continuous variable within each smoking category.

```{r}
#| echo: false


df |> 
  tbl_summary(
    by = low,
  type = list(bwt ~ "continuous", lwt ~ "continuous", age ~ "continuous"),
  statistic = all_continuous() ~ "{median} ({IQR})",
  label = list(age ~ "Age of Mother", 
               lwt ~ "Weight of Mother",
               bwt ~ "Birth Weight of the, Child",
               race ~ "Race")
)
```

#### Answer

```{r}
#| code-fold: true
#| eval: false


df |> 
  tbl_summary(
    by = low,
  type = list(bwt ~ "continuous", lwt ~ "continuous", age ~ "continuous"),
  statistic = all_continuous() ~ "{median} ({IQR})",
  label = list(age ~ "Age of Mother", 
               lwt ~ "Weight of Mother",
               bwt ~ "Birth Weight of the, Child",
               race ~ "Race")
)
```

### **Question 5:** Linear Regression Model

- Fit a linear regression model to predict birth weight (`bwt`) using the mother's weight (`lwt`) and race (`race`).
- Use the `tbl_regression()` function to display the model summary, including coefficients and p-values.

```{r}
#| echo: false

# Question 5: Linear Regression Model
model <- lm(bwt ~ lwt + race, data = df)

model |> 
tbl_regression()
```

#### Answer

```{r}
#| code-fold: true
#| eval: false

# Question 5: Linear Regression Model
model <- lm(bwt ~ lwt + race, data = df)

model |> 
tbl_regression()
```

